Tool-driven Revolutions in Archaeological Science

### 2018-09-10 analysis/paper/jcaa-paper.Rmd
add Stodden et al - reference
typo

There is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses. In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences.

@@

The result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices in scientific computing (described in more detail in @marwick2018packaging) is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools). The goal of rrtools ist o make it easier for archaeologists and researchers in other fields to use R for research and publication. This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort. The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research. First, we keep the data separate from the methods. This means keep the code in a separate directory to the data. Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.

>>>>

There is a longstanding tradition in archaeology of publishing catalogues, lists and maps showing the data used for analyses. In the digital age, though, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences. These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013].

@@

The result of our analysis of existing practices among archaeologists and other researchers, and our study of best practices in scientific computing (described in more detail in @marwick2018packaging) is the R package rrtools ('reproducible research tools', https://github.com/benmarwick/rrtools). The goal of rrtools ist to make it easier for archaeologists and researchers in other fields to use R for research and publication. This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort. The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research. First, we keep the data separate from the methods. This means keep the code in a separate directory to the data. Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
